---
output:
  html_document:
    toc:           False
    theme:         cerulean
    highlight:     pygments
   
---

### Nesting Versus Piping

R is primarily a functional programing language, which leads to a lot of nesting:   `thenDoSomeThingElse(doSomethingFirst(data))`.   Notice that we start from within  with orginal data and then move out.  Even a moderate amount of nesting can make your code difficult to read, including for your future self.  That’s why I have been absolutely in enthralled with the new piping operator `%>%` from the  `magrittr` package and used heavily in the data manpulation package  `dplyr`. 

The piping operator passes an object as the first agrument into the following function.  Going back our earlier example: `data %>% doSomethingFirst() %>% thenDoSomeThingElse()`. Piping allows code in a much more natural way, from left to right.

### Piping Example With dplyr
Lets go through a more expansive example using flight data 2013 New York City flight data.   Now, let’s say that we are interested in the following question: What airlines had the most flights headed for O’Hare Airport (ORD) on the first day of the month?

We’ll answer this question using dplyr functions using traditional nesting and then using the piping operator.  Check out the full code on github. 

First, let's take a peak at the data. 

```{r,echo=FALSE, message=FALSE}

library(dplyr);library(nycflights13)

data(flights)

``` 

```{r,comment=NA}
flights


```

Looks like a lot of good data.  Now let's answer our question using nesting.

```{r,comment=NA}

flightsToChicago_NESTING<-arrange(
  summarize(
    group_by(
      filter(flights,dest=="ORD",day==1),carrier),numFlights=n()),desc(numFlights)
  )

flightsToChicago_NESTING


```

Alright, looks like United Airlines comes up on top.  We started by filtering the  `flights` data to observations where the destination`dest` is O'hare and the day of the month `day` is one, but that can be hard to miss because of all the nesting business going on.

Now, let's consider piping. 

```{r,comment=NA}


flightsToChicago_PIPING<-flights%>%
                            filter(dest=="ORD",day==1)%>%
                            group_by(carrier)%>%
                            summarise(numFlights=n())%>%
                            arrange(desc(numFlights))

flightsToChicago_PIPING


```

Boom! More readable code.  Bottomline: piping makes data analysis more enjoyable.



