---
output:
  html_document:
    toc:           False
    theme:         cerulean
    highlight:     pygments
   
---

### Nesting Versus Piping

R is primarily a functional programing language, which leads to a lot of nesting:   `thenDoSomethingElse(doSomethingFirst(data))`.   Notice that we start from within  with the original data and then move out.  Even a moderate amount of nesting can make your code difficult to read, including for your future self!  That’s why I've been absolutely enthralled with the new piping operator `%>%` from the  `magrittr` package.  The operator works well in concert with Hadley Wickham's excellent new data wrangling package  `dplyr`. 

The piping operator passes an object as the first argument into the following function.  Going back our earlier example: `data %>% doSomethingFirst() %>% thenDoSomeThingElse()`.  You might think of the operator as equivanant to "then do this".  Piping allows code to be read more naturally from left to right.

</br>

### Piping Example with dplyr

Lets go through a more expansive example using 2013 New York City flight data.   Now, let’s say that we are interested in the following question: **What airlines had the most flights headed for O’Hare Airport (ORD) on the first day of the month?**

We’ll answer this question using dplyr functions first using traditional nesting and then using the piping operator.  Check out the full code on [github](https://github.com/mannybernabe/pipingExample). 

First, let's take a peak at the data. 

```{r,echo=FALSE, message=FALSE}

library(dplyr);library(nycflights13)

data(flights)

``` 

```{r,comment=NA}
flights


```

Looks like a lot of good data.  Now let's answer our question using nesting.

```{r,comment=NA}

flightsToChicago_NESTING<-arrange(
  summarize(
    group_by(
      filter(flights,dest=="ORD",day==1),
      carrier),
    numFlights=n()),
  desc(numFlights)
)

flightsToChicago_NESTING


```

Alright, looks like United Airlines comes up on top.  We started by filtering the  `flights` data to observations where the destination(`dest`) equals O'hare and the day of the month (`day`) equals one, but that can be hard to miss with all the nesting business going on.

Now, let's consider piping. 

```{r,comment=NA}


flightsToChicago_PIPING<-flights%>%
                            filter(dest=="ORD",day==1)%>%
                            group_by(carrier)%>%
                            summarise(numFlights=n())%>%
                            arrange(desc(numFlights))

flightsToChicago_PIPING


```

Boom! More readable code.  Bottomline: piping makes data analysis more enjoyable.



